---
title: "Selecting Community Detection Resolution Parameters with CHAMP Maps: the football network"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Selecting Community Detection Resolution Parameters with CHAMP Maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 3,
  fig.width = 5
)
```

```{r setup}
library(ideanet)
```

**This vignette takes a deeper dive into using the framework of CHAMP and maps on the CHAMP set to find communities aligned with the conference structure in the network of Division I-A college football games from the Fall 2000 season. If you haven't already done so, you will want to look at the `CHAMP` vignette to understand the commands used here.**

These functions, as part of the broader IDEANet project, are supported by the National Science Foundation as part of the Human Networks and Data Science - Infrastructure program (BCS-2024271 and BCS-2140024).

If you use these functions in your work, please cite

1.  [the `ideanet` package](https://doi.org/10.32614/CRAN.package.ideanet),

2.  [the Weir et al. (2017) CHAMP paper](https://doi.org/10.3390/a10030093), and

3.  [the Gibson & Mucha (2022) CHAMP maps paper](https://doi.org/10.1038/s41598-022-20142-6).

## Example: NCAA College Football Network

A favorite example for demonstrating the need for a resolution parameter is the network of the (then-so-called) Division I-A college football games from the Fall 2000 season. This example was introduced by Girvan and Newman (2002). The GML file for this example can be downloaded (along with a bunch of other interesting examples) from Mark Newman’s “Network data” web page at <http://www-personal.umich.edu/~mejn/netdata/>. We have directly included this `football` data in `ideanet` with permission from Girvan and Newman.

```{r, eval = FALSE}
football
```

```{r football_viz, eval = FALSE}
igraph::plot.igraph(football, main="Div-IA College Football, Fall 2000")
```

We can now run the three steps of this framework on the football data as follows.

```{r, warning = FALSE, eval = FALSE}
football_partitions <- get_partitions(football, n_runs = 500, seed = 115)
football_partitions <- CHAMP(football, football_partitions,
                             plottitle="Div-IA College Football, Fall 2000")
football_partitions <- get_CHAMP_map(football, football_partitions,
                                     plotlabel="Div-IA College Football, Fall 2000")
```

But if we were to run the commands using default parameters like above, we would find that the fixed point of the iterative map on the CHAMP set is the partition that is optimal at the largest values of $\gamma$ in the default range considered. (You can see this either by running the above manually, or you will see it in what follows now.) This is a sign to us that we might want to consider even larger values of $\gamma$, which we can easily do by changing `gamma_range` from its default in the call to `get_partitions` as follows. Note that this changes the resulting set of `football_partitions` obtained, which then get passed to `CHAMP` and `get_CHAMP_maps`.

```{r, warning = FALSE}
football_partitions <- get_partitions(football, n_runs = 500, 
                                      gamma_range = c(0,7), seed = 115)
football_partitions <- CHAMP(football, football_partitions,
                             plottitle="Div-IA College Football, Fall 2000")
football_partitions <- get_CHAMP_map(football, football_partitions,
                                     plotlabel="Div-IA College Football, Fall 2000")
print(football_partitions$CHAMPsummary)
```

Now we see that everything in the figure above flows towards the 12-community partition (`partition_num` 27) from both below and above, and we emphasize that the range of $\gamma$ for which this is the optimal partition does not straddle the default $\gamma=1$ value. While this 12-community partition fixed point appears to be the same fixed point partition we found above with the restricted default `gamma_range`, it is only because we extended the range of $\gamma$ considered that we can see that the maps point towards this fixed point from both above and below. It can be important in developing your CHAMP and iterative map results that you consider $\gamma$ values sufficiently above and below the fixed points that you find.

As with the karate club, we highlight that the result here is different from that found in Newman (2016), where the number of communities is pre-specified to be 11 for this network. In contrast, by allowing for different possible numbers of communities, the fixed point identified here has 12 communities. We also note that extending the `gamma_range` beyond its default here identified a 14-community partition (`partition_num` 31) that is the optimum over a wide range of $\gamma$. While this partition is not a fixed point — indeed, it points to the 12-community partition — it might be interesting in practice to further consider such a partition, since it has one of the largest domains of optimality in gamma. (We will briefly return to both the 11-community and the 14-community partitions below.)

### What are the communities finding here?

To finish this example, we look more closely at these communities we found to demonstrate that they are useful for understanding this data set. That is, having used the CHAMP and iterative map framework to find only a single fixed-point partition, and to have it not be at the default $\gamma=1$ resolution, we ask ourselves, what are these communities in the fixed-point partition?

```{r football_partitions_viz, eval = FALSE}
colrs <- c("#543005", "#8C510A", "#BF812D", "#DFC27D", "#F6E8C3", "#C7EAE5", "#80CDC1", "#35978F", "#01665E", "#003C30")
igraph::V(football)$color <- colrs[igraph::V(football)$value+1]
igraph::plot.igraph(football,mark.groups = football_partitions$partitions[[27]],
                    vertex.label=NA, vertex.size=5)
```

```{r football_table}
table(igraph::V(football)$value,
      igraph::membership(football_partitions$partitions[[27]]))
```

Though it isn't completely clear from the figure, the above table indicates that this 12-community fixed-point partition is very well aligned to the `value` labels in the data. This similarity between labels is natural because `value` here denotes the conferences that teams participated in (see <https://en.wikipedia.org/wiki/Category:2000_NCAA_Division_I-A_football_standings_templates>). For example, we see in the table that Community 1 consists of all 8 teams with `value=7` (this value corresponds to the old Mountain West Conference). Continuing through the table, Community 2 is all 9 teams with `value=0` (the ACC), Community 3 is all 11 teams with `value=2` (the Big Ten), Community 4 is all 12 teams with `value=3` (the Big 12), Community 5 is all 10 teams with `value=8` (the Pac-10) and Community 8 is all 12 teams with `value=9` (the SEC).

Looking at a community here that is not perfectly aligned with the `value` attribute, we find Community 7

```{r community7}
igraph::V(football)$label[football_partitions$partitions[[27]]$membership == 7]
```

is the 13 teams of the Mid-American Conference (`value=6`) along with independent Connecticut, who played a number of games against these teams:

```{r uconn}
UConn_nbhd <- unlist(igraph::neighborhood(football, 
                        nodes = which(igraph::V(football)$label=="Connecticut")))
igraph::V(football)$label[UConn_nbhd]
```

Note here that the igraph neighborhood includes the selected node (Connecticut). Note also that Connecticut played more than 7 games in 2000, but the other teams are not included in this data set because they were not Division I-A teams.

As one more interesting look at these communities, we see that Community 11

```{r community11}
igraph::V(football)$label[football_partitions$partitions[[27]]$membership == 11]
```

aligns perfectly with the 9 teams in Conference USA, while Community 12

```{r community12}
igraph::V(football)$label[football_partitions$partitions[[27]]$membership == 12]
```

aligns perfectly with the 9 teams in the Western Athletic Conference. The apparent discrepancy between these memberships and the `value` field is that TexasChristian (TCU) appears to be misassigned by the `value` field to Conference USA (`value=4`) instead of the WAC (`value=11`), which also erroneously appears to include BoiseState from the Big West and independent LouisianaTech.

Finally, as one last emphasis that sometimes partitions that are not fixed points are still of potential interest, and having highlighted the large-gamma-range `partition_num` 31, we compare this 14-community partition to the 12-community fixed point:

```{r table36_41}
table(igraph::membership(football_partitions$partitions[[27]]),
      igraph::membership(football_partitions$partitions[[31]]))
```

You can see from this table that these two partitions are identical except for the differences in the splitting of the 12-community fixed point partition's "Community 7" 14-node community into two 7-node communities, and breaking out two teams from "Community 9". If you query these differences by team name, the splitting of Community 7 into two matches the split of the Mid-American Conference into its East and West divisions (grouping independent Connecticut in the West), while the two teams broken off from "Community 9" are independents Navy and NotreDame from the 8 teams of the Big East Conference.

Similarly, we again note that Newman (2016) specified 11 communities. Our results above found an 11-community solution, `partition_num` 24, optimal for $\gamma$ between 1.20 and 1.45, that maps to a $\gamma$ estimate of 2.27 (where the 12-community partition is optimal). Since this estimated $\gamma$ matches the value in Table I of Newman (2016), we believe it is likely the same partition found there and we compare it with the 12-community fixed point:

```{r table36_33}
table(igraph::membership(football_partitions$partitions[[27]]),
      igraph::membership(football_partitions$partitions[[24]]))
```

You can see from the above table that the two partitions are almost identical, with both of them providing excellent identification of the conference structure. The only differences all involve the 5 teams in Community 10 of the 12-community partition,

```{r}
igraph::V(football)$label[football_partitions$partitions[[27]]$membership == 10]
```

which, importantly, are all independents (along with Connecticut, Navy and Notre Dame). Querying further, it turns out that the 11-community partition groups independent CentralFlorida (UCF) in with the Mid-American Conference (plus Connecticut, as discussed previously) and puts these other four listed teams (all independents) with Conference USA. Is one of these solutions "better" than the other? Not necessarily. It is on the one hand a nice feature of the 12-community partition that these 5 independents are not assigned to any of the conference-aligned communities. But on the other hand, these assignments of these independent teams are not unreasonable given the scheduling quirks of independent teams.

In summary, we find that this 12-community fixed point partition is an excellent clustering of this network data set, with important qualities that are much better than what we would have obtained at the default resolution value ($\gamma=1$).

## References

Girvan, M., and M. E. J. Newman. “Community Structure in Social and Biological Networks.” Proceedings of the National Academy of Sciences of the United States of America 99, no. 12 (June 11, 2002): 7821–26. <https://doi.org/10.1073/pnas.122653799>.

Gibson, Ryan A., and Peter J. Mucha. 2022. "Finite-State Parameter Space Maps for Pruning Partitions in Modularity-Based Community Detection." Scientific Reports 12 (1): 15928. <https://doi.org/10.1038/s41598-022-20142-6>.

-   Python implementation: Gibson, Ryan. 2020--2024. <https://github.com/ragibson/ModularityPruning>.

Newman, M. E. J. “Equivalence between Modularity Optimization and Maximum Likelihood Methods for Community Detection.” Physical Review E 94, no. 5 (November 22, 2016): 052315. <https://doi.org/10.1103/PhysRevE.94.052315>.

Pamfil, A. Roxana., Sam D. Howison, Renaud. Lambiotte, and Mason A. Porter. “Relating Modularity Maximization and Stochastic Block Models in Multilayer Networks.” SIAM Journal on Mathematics of Data Science, January 1, 2019, 667–98. <https://doi.org/10.1137/18M1231304>.

Peel, Leto, Daniel B. Larremore, and Aaron Clauset. 2017. "The Ground Truth about Metadata and Community Detection in Networks." Science Advances 3 (5): e1602548. <https://doi.org/10.1126/sciadv.1602548>.

Weir, William H., Scott Emmons, Ryan Gibson, Dane Taylor, and Peter J. Mucha. 2017. "Post-Processing Partitions to Identify Domains of Modularity Optimization." Algorithms 10 (3): 93. <https://doi.org/10.3390/a10030093>.

-   Python implementation: Weir, William. 2017--2018. <https://github.com/wweir827/CHAMP>.
